---
title: "More about Quarto"
format:
  html:
    code-fold: true
bibliography: references.bib
toc: true
number-sections: true
---

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
```

```{r}
#| label: load-data
#| include: false
penguins <- read_csv("../data/penguins.csv")
```

# Computations

Code and its output are both displayed in the rendered document by default.

```{r}
#| label: fig-scatterplot
#| fig-cap: "Body mass and bill length for penguins"
#| fig-alt: "Scatterplot of body mass vs. bill length for penguins. The plot displays a positive relationship between body mass and bill length."
#| fig-width: 6
#| fig-height: 3

ggplot(penguins) + 
  aes(x = body_mass_g, y = bill_length_mm) + 
  geom_point()
```

You can tweak the defaults by adding to the YAML header at the very top. Let's try this example: for some documents, you might want to hide all the code.

--\> Go to the YAML header and specify (copy and paste):

```         
execute:
  echo: false
```

(Note the two spaces before echo.)

Render the file.

You can also override the document level settings. For example, you might want to selectively render some code chunks.

--\> Go to the fig-scatterplot chunk, and add `#| echo: true`. Render. The code is now included in the fig-scatterplot chunk.

## Code Folding

You can allow people to view the code only if they want to. Try the `code-fold` option.

--\> Remove the `execute: echo` option. Add in the `code-fold` option under `html`, like so:

```         
---
title: "More about Quarto"
format:
  html:
    code-fold: true
---
```

--\> Render/save again, and see the Code widgets for each chunk.

## Figures

Quarto gives you the ability to improve the appearance and accessibility of your plots. We can change the size of our plot with `fit-width` and `fig-height`, provide a caption with `fig-cap` , add alt text for accessibility with `fig-alt`.

--\> Copy and paste over the existing code-block options into the plot above.

```         
#| label: fig-scatterplot
#| fig-cap: "Body mass and bill length for penguins"
#| fig-alt: "Scatterplot of body mass vs. bill length for penguins. The plot displays a positive relationship between body mass and bill length."
#| fig-width: 6
#| fig-height: 3
```

You can reference the plot in the text:

@fig-scatterplot displays a positive relationship between body mass and bill length.

## Inline Code

You can also include inline code (code not inside a code chunk) by surrounding your code with backticks and prefacing the code with r.

--\> Include inline code below. Replace the ... with your code.

--\> Code: nrow(penguins). Include backticks and preface the code with r. You can also use the dropdown: Format \> Code for the backticks.

There are `r nrow(penguins)` penguins in the dataset.

--\> Render.

The sentence will display the calculated number of observations.

If your code is a bit longer, you can also create an object in a code chunk with `echo: false`, and call that object in the inline code.

```{r}
#| echo: false

mean_bill <- round(mean(penguins$bill_length_mm, na.rm=TRUE), 2)
```

The average bill length of the penguins in our data is `r mean_bill`.

## Caching

If you have code chunks that will take a long time to compute, you can cache the results. You can do this globally with a YAML header:

```         
execute:
  cache: true
```

Or you can cache individual code chunks:

```         
#| cache: true
```

## Equations

You can add in equations using LaTeX math.

--\> Go to Insert in the toolbar, LaTeX Math, then Display Math. Type in y = x\^2.

$$
y = x^2
$$

## Citations

It is easy to include citations in Quarto. You can either connect Quarto to Zotero, import a .bibtex file that has a list of citations, import citations by their DOI, and a few other methods.

--\> Go to Insert in the Toolbar, Select Citation, Select **From DOI** on the left and copy-and-paste the DOI `10.5281/zenodo.3960218` in the search bar and hit **Search**. Then, select the found reference, and **Insert** it into your document, below (before the period). Render.

We are using the Palmer Penguins dataset [@horst2020].

Notice the YAML header added:

```         
#| bibliography: references.bib
```

And we now have the *references.bib* file in our project directory.

## Table of Contents and Themes

We can do a lot to fine tune our document. There are so many options here that we can't cover them all. Let's just do just a couple of simple changes to our document. We will add a Table of Contents and change the Theme.

--\> Add this to the YAML header.

```         
toc: true
number-sections: true
```

--\>Now add a theme to change the look and style of the document. Find more [themes](https://quarto.org/docs/output-formats/html-themes.html).

```         
format:
  html:
    theme: quartz
```

--\> Render.

## Presentations and PDFs

Finally, let's look at how to make Quarto documents from scratch, including Presentations and PDFs.

--\> Go to File...New File...Quarto Presentation.

--\> Go to File...New File...Quarto Document...PDF.

Note that PDF output requires a LaTeX installation, like tinytex.

You can install with the following command, in the Terminal.

```         
quarto install tool tinytex
```
